\bf A Comparision of Bootstrap Confidence Interval Methods

### 2022-12-07 project/PaperTemplate/paper.tex
Rewrote beginning of simulation write-up

  \author{
        Justin Papagelis \thanks{The authors gratefully acknowledge
\ldots{}} \\
    Department of Mathematics and Statistics, Amherst College\\
      }
  \maketitle
} \fi

@@

The standard confidence interval is built based on the assumption that
the distribution from which we are sampling is Normal. This means that
for an unknown distribution, the standard confidence interval could
present an incorrect range. However, the same process can be used with
bootstrap sampling to form the bootstrap percentile method. This means
that an approximate bootstrap confidence interval will be created in the
same automatic way that the standard confidence interval was created.
For bootstrapped confidence intervals, the number of bootstrap
replications \(B\) must be large (around 2000) due to the nature of
confidence intervals requiring greater accuracy \citep{Efron86}.

@@

For the simulation, we will randomly sample from distributions in which
we know the true population mean. For example, we will sampling from a
Gamma Distribution where we can find the true population mean using the
shape and rate parameters of the sample distribution. After we sample
from a distribution, we will create a bootstrapped confidence interval
for each of the methods we described in the exposition. Since we know
the true population mean of the distribution, we will determine if the
true population parameter is contained in the confidence interval for
each of the methods. We will perform this process a large number of
times (10,000) and then average by the number of iterations to determine
the average proportion that each bootstrapped confidence interval method
contains the true population mean. We will do this for a couple
different distributions as well as a range of sample sizes show how the
intervals perform with skewed distributions and larger sample sizes.

First, we create a function that calculates the bootstrapped confidence
intervals from a given sample. To do this, we first bootstrap the sample
and then perform each method.

Next, we will create the simulation that picks the random sample from a
distribution and tests to see if the sample creates bootstrapped
confidence intervals that contain the true population parameter. First,
we have the simulation of selecting from a Normal Distribution. Since we
have a \(N(0,1)\) distribution, out true population mean is 0.

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(boot)}

\NormalTok{run\_sim\_log\_normal }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(sim\_reps) \{}
  \FunctionTok{set.seed}\NormalTok{(}\DecValTok{495}\NormalTok{)}
\NormalTok{  mean.fun }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(d, i) \{}
\NormalTok{    d }\OtherTok{\textless{}{-}}\NormalTok{ X[i]}
    \FunctionTok{return}\NormalTok{ (}\FunctionTok{c}\NormalTok{(}\FunctionTok{mean}\NormalTok{(d), }\FunctionTok{sd}\NormalTok{(d)))}
\NormalTok{  \}}

@@

\NormalTok{      results }\OtherTok{\textless{}{-}} \FunctionTok{boot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ values, }\AttributeTok{statistic =}\NormalTok{ mean.fun, }\AttributeTok{R =}\NormalTok{ boot\_reps)}
\NormalTok{      results}
      \CommentTok{\#plot(results)}
\NormalTok{      ci }\OtherTok{\textless{}{-}} \FunctionTok{boot.ci}\NormalTok{(results, }\AttributeTok{type =} \StringTok{"all"}\NormalTok{)}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(boot)}

\NormalTok{run\_sim\_normal }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(sim\_reps) \{}
  \FunctionTok{set.seed}\NormalTok{(}\DecValTok{495}\NormalTok{)}
\NormalTok{  mean.fun }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(d, i) \{}
\NormalTok{    d }\OtherTok{\textless{}{-}}\NormalTok{ X[i]}
    \FunctionTok{return}\NormalTok{ (}\FunctionTok{c}\NormalTok{(}\FunctionTok{mean}\NormalTok{(d), }\FunctionTok{sd}\NormalTok{(d)))}
\NormalTok{  \}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(boot)}

\NormalTok{run\_sim\_gamma }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(sim\_reps) \{}
  \FunctionTok{set.seed}\NormalTok{(}\DecValTok{495}\NormalTok{)}
\NormalTok{  mean.fun }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(d, i) \{}
\NormalTok{    d }\OtherTok{\textless{}{-}}\NormalTok{ X[i]}
    \FunctionTok{return}\NormalTok{ (}\FunctionTok{c}\NormalTok{(}\FunctionTok{mean}\NormalTok{(d), }\FunctionTok{sd}\NormalTok{(d)))}
\NormalTok{  \}}

@@

\NormalTok{gamma\_df\_long }\OtherTok{\textless{}{-}}\NormalTok{ gamma\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \DecValTok{2}\SpecialCharTok{:}\DecValTok{6}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"method"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"prop"}\NormalTok{) }

@@

\NormalTok{gamma\_table }\OtherTok{\textless{}{-}}\NormalTok{ gamma\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable}\NormalTok{(}\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{align =} \StringTok{"c"}\NormalTok{, }\AttributeTok{caption =} \StringTok{"Sampled From Gamma Dist."}\NormalTok{, }
        \AttributeTok{col.names =} \FunctionTok{gsub}\NormalTok{(}\StringTok{"[\_]"}\NormalTok{, }\StringTok{" "}\NormalTok{, }\FunctionTok{names}\NormalTok{(gamma\_df))) }

\NormalTok{normal\_df\_long }\OtherTok{\textless{}{-}}\NormalTok{ normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \DecValTok{2}\SpecialCharTok{:}\DecValTok{6}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"method"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"prop"}\NormalTok{) }

@@

\NormalTok{normal\_table }\OtherTok{\textless{}{-}}\NormalTok{ normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable}\NormalTok{(}\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{align =} \StringTok{"c"}\NormalTok{, }\AttributeTok{caption =} \StringTok{"Sampled from Normal Dist."}\NormalTok{, }
        \AttributeTok{col.names =} \FunctionTok{gsub}\NormalTok{(}\StringTok{"[\_]"}\NormalTok{, }\StringTok{" "}\NormalTok{, }\FunctionTok{names}\NormalTok{(normal\_df)))}

\NormalTok{log\_normal\_df\_long }\OtherTok{\textless{}{-}}\NormalTok{ log\_normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \DecValTok{2}\SpecialCharTok{:}\DecValTok{6}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"method"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"prop"}\NormalTok{) }

@@

\NormalTok{log\_normal\_table }\OtherTok{\textless{}{-}}\NormalTok{ log\_normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable}\NormalTok{(}\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{align =} \StringTok{"c"}\NormalTok{, }\AttributeTok{caption =} \StringTok{"Sampled from Normal Dist."}\NormalTok{, }
        \AttributeTok{col.names =} \FunctionTok{gsub}\NormalTok{(}\StringTok{"[\_]"}\NormalTok{, }\StringTok{" "}\NormalTok{, }\FunctionTok{names}\NormalTok{(log\_normal\_df)))}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-24-1.pdf}

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-24-2.pdf}

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-24-3.pdf}

@@

\caption{\label{tab:create graphs}Sampled From Gamma Dist.}
\centering
\begin{tabular}[t]{cccccc}
\toprule
Sample Size & Normal & Basic & Studentized & Percentile & BCa\\
\midrule
4 & 0.8494 & 0.8332 & 0.9039 & 0.8578 & 0.8750\\
10 & 0.8476 & 0.8291 & 0.9008 & 0.8543 & 0.8708\\
15 & 0.8510 & 0.8322 & 0.9012 & 0.8563 & 0.8706\\
20 & 0.8458 & 0.8274 & 0.9030 & 0.8511 & 0.8672\\
30 & 0.8380 & 0.8218 & 0.8954 & 0.8463 & 0.8642\\
\addlinespace
40 & 0.8480 & 0.8313 & 0.8989 & 0.8525 & 0.8664\\
60 & 0.8466 & 0.8292 & 0.9069 & 0.8571 & 0.8741\\
80 & 0.8486 & 0.8310 & 0.9016 & 0.8535 & 0.8717\\
100 & 0.8478 & 0.8308 & 0.8974 & 0.8533 & 0.8660\\
\bottomrule
\end{tabular}
\end{table}

@@

\NormalTok{master\_fig }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ master\_df\_long, }
                    \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Sample\_Size, }\AttributeTok{y =}\NormalTok{ prop, }\AttributeTok{color =}\NormalTok{ method)) }\SpecialCharTok{+} 
  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+} \FunctionTok{geom\_line}\NormalTok{() }\SpecialCharTok{+} 
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Proportion of CIs that Contain the True Parameter from Log{-}Normal}
\StringTok{       Distibution"}\NormalTok{, }\AttributeTok{x =} \StringTok{"Sample Size"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Proportion (of 10,000 Simulations)"}\NormalTok{, }\AttributeTok{color =} \StringTok{"Method"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 8: Graph showing the Accuracy of Bootstrap CI Methods for different sample sizes on Log{-}Normal Dist."}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_light}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position=}\StringTok{"bottom"}\NormalTok{)}

>>>>

  \author{
        Justin Papagelis \thanks{The authors gratefully acknowledge
Profossor Wagaman} \\
    Department of Mathematics and Statistics, Amherst College\\
      }
  \maketitle
} \fi

@@

The standard confidence interval is built based on the assumption that
the distribution from which we are sampling is Normal. As such, the
standard confidence interval is sometimes called the normal confidence
interval for bootstrapped data. This means that for an unknown
distribution, the standard confidence interval could present an
incorrect range. However, the same process can be used with bootstrap
sampling to form the bootstrap percentile method. This means that an
approximate bootstrap confidence interval will be created in the same
automatic way that the standard confidence interval was created. For
bootstrapped confidence intervals, the number of bootstrap replications
\(B\) must be large (around 2000) due to the nature of confidence
intervals requiring greater accuracy \citep{Efron86}.

@@

For the simulation, we will randomly sample from distributions in which
we know the true population mean. For example, we will sampling from a
Gamma Distribution where we can find the true population mean using the
shape and rate parameters of the sample distribution. After we sample
from a distribution, we will create a bootstrapped confidence interval
for a few of the methods we described in the exposition. Since we know
the true population mean of the distribution, we will determine if the
true population parameter is contained in the confidence interval for
each of the methods we are testing. We will perform this process a large
number of times (10,000 simulations) and then average by the number of
iterations to determines on average, the proportion that each
bootstrapped confidence interval method contains the true population
mean for a given distribution. We will do this for a couple different
distributions as well as a range of sample sizes show how the intervals
perform with skewed distributions and larger sample sizes.

To do this simulation, we will be using the \texttt{boot} package
\citep{boot}. This package can compute a bootstrap object which contains
the output from a bootstrap calculation of a given sample and then
produces confidence intervals from the bootstrap object. In particular,
we will use the \texttt{boot.ci} function to create four types of
confidence intervals: the standard (or normal) confidence interval, the
studentized confidence interval, the percentile confidence interval, and
the bias-corrected with acceleration confidence interval.

First, we will create a simulation that pick a random sample from a
distribution and performs

and tests to see if the sample creates a bootstrapped confidence
intervals that contain the true population parameter. First, we have the
simulation of selecting from a Normal Distribution. Since we have a
\(N(0,1)\) distribution, out true population mean is 0.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{run\_sim\_log\_normal }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(sim\_reps) \{}
  \FunctionTok{set.seed}\NormalTok{(}\DecValTok{495}\NormalTok{)}
\NormalTok{  mean.fun }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(d, i) \{}
\NormalTok{    d }\OtherTok{\textless{}{-}}\NormalTok{ X[i]}
    \FunctionTok{return}\NormalTok{ (}\FunctionTok{c}\NormalTok{(}\FunctionTok{mean}\NormalTok{(d), }\FunctionTok{sd}\NormalTok{(d)))}
\NormalTok{  \}}

@@

\NormalTok{      results }\OtherTok{\textless{}{-}} \FunctionTok{boot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ values, }\AttributeTok{statistic =}\NormalTok{ mean.fun, }\AttributeTok{R =}\NormalTok{ boot\_reps)}
\NormalTok{      results}
      \CommentTok{\#plot(results)}
\NormalTok{      ci }\OtherTok{\textless{}{-}} \FunctionTok{boot.ci}\NormalTok{(results, }\AttributeTok{conf =} \FloatTok{0.95}\NormalTok{, }\AttributeTok{type =} \FunctionTok{c}\NormalTok{(}\StringTok{"norm"}\NormalTok{, }\StringTok{"perc"}\NormalTok{, }\StringTok{"bca"}\NormalTok{, }\StringTok{"stud"}\NormalTok{))}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{run\_sim\_normal }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(sim\_reps) \{}
  \FunctionTok{set.seed}\NormalTok{(}\DecValTok{495}\NormalTok{)}
\NormalTok{  mean.fun }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(d, i) \{}
\NormalTok{    d }\OtherTok{\textless{}{-}}\NormalTok{ X[i]}
    \FunctionTok{return}\NormalTok{ (}\FunctionTok{c}\NormalTok{(}\FunctionTok{mean}\NormalTok{(d), }\FunctionTok{sd}\NormalTok{(d)))}
\NormalTok{  \}}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{run\_sim\_gamma }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(sim\_reps) \{}
  \FunctionTok{set.seed}\NormalTok{(}\DecValTok{495}\NormalTok{)}
\NormalTok{  mean.fun }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(d, i) \{}
\NormalTok{    d }\OtherTok{\textless{}{-}}\NormalTok{ X[i]}
    \FunctionTok{return}\NormalTok{ (}\FunctionTok{c}\NormalTok{(}\FunctionTok{mean}\NormalTok{(d), }\FunctionTok{sd}\NormalTok{(d)))}
\NormalTok{  \}}

@@

\CommentTok{\# create the figure that shows the gamma output}
\NormalTok{gamma\_df\_long }\OtherTok{\textless{}{-}}\NormalTok{ gamma\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \DecValTok{2}\SpecialCharTok{:}\DecValTok{6}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"method"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"prop"}\NormalTok{) }

@@

\CommentTok{\# create the table for the gamma output}
\NormalTok{gamma\_table }\OtherTok{\textless{}{-}}\NormalTok{ gamma\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable}\NormalTok{(}\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{align =} \StringTok{"c"}\NormalTok{, }\AttributeTok{caption =} \StringTok{"Sampled From Gamma Dist."}\NormalTok{, }
        \AttributeTok{col.names =} \FunctionTok{gsub}\NormalTok{(}\StringTok{"[\_]"}\NormalTok{, }\StringTok{" "}\NormalTok{, }\FunctionTok{names}\NormalTok{(gamma\_df)), }\AttributeTok{digits =} \DecValTok{3}\NormalTok{) }

\CommentTok{\# create the figure that shows the normal output }
\NormalTok{normal\_df\_long }\OtherTok{\textless{}{-}}\NormalTok{ normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \DecValTok{2}\SpecialCharTok{:}\DecValTok{6}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"method"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"prop"}\NormalTok{) }

@@

\CommentTok{\# create the table for the normal output}
\NormalTok{normal\_table }\OtherTok{\textless{}{-}}\NormalTok{ normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable}\NormalTok{(}\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{align =} \StringTok{"c"}\NormalTok{, }\AttributeTok{caption =} \StringTok{"Sampled from Normal Dist."}\NormalTok{, }
        \AttributeTok{col.names =} \FunctionTok{gsub}\NormalTok{(}\StringTok{"[\_]"}\NormalTok{, }\StringTok{" "}\NormalTok{, }\FunctionTok{names}\NormalTok{(normal\_df)))}

\CommentTok{\# create the figure that shows the log normal output}
\NormalTok{log\_normal\_df\_long }\OtherTok{\textless{}{-}}\NormalTok{ log\_normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{pivot\_longer}\NormalTok{(}\AttributeTok{cols =} \DecValTok{2}\SpecialCharTok{:}\DecValTok{6}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"method"}\NormalTok{, }\AttributeTok{values\_to =} \StringTok{"prop"}\NormalTok{) }

@@

\CommentTok{\# create the table for the log normal output}
\NormalTok{log\_normal\_table }\OtherTok{\textless{}{-}}\NormalTok{ log\_normal\_df }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{kable}\NormalTok{(}\AttributeTok{booktabs =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{align =} \StringTok{"c"}\NormalTok{, }\AttributeTok{caption =} \StringTok{"Sampled from Normal Dist."}\NormalTok{, }
        \AttributeTok{col.names =} \FunctionTok{gsub}\NormalTok{(}\StringTok{"[\_]"}\NormalTok{, }\StringTok{" "}\NormalTok{, }\FunctionTok{names}\NormalTok{(log\_normal\_df)))}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{paper_files/figure-latex/output the graphs/tables-1.pdf}

@@

\includegraphics{paper_files/figure-latex/output the graphs/tables-2.pdf}

@@

\includegraphics{paper_files/figure-latex/output the graphs/tables-3.pdf}

@@

\caption{\label{tab:create graphs}Sampled From Gamma Dist.}
\centering
\begin{tabular}[t]{cccccc}
\toprule
Sample Size & Normal & Basic & Studentized & Percentile & BCa\\
\midrule
4 & 0.849 & 0.833 & 0.904 & 0.858 & 0.875\\
10 & 0.848 & 0.829 & 0.901 & 0.854 & 0.871\\
15 & 0.851 & 0.832 & 0.901 & 0.856 & 0.871\\
20 & 0.846 & 0.827 & 0.903 & 0.851 & 0.867\\
30 & 0.838 & 0.822 & 0.895 & 0.846 & 0.864\\
\addlinespace
40 & 0.848 & 0.831 & 0.899 & 0.853 & 0.866\\
60 & 0.847 & 0.829 & 0.907 & 0.857 & 0.874\\
80 & 0.849 & 0.831 & 0.902 & 0.854 & 0.872\\
100 & 0.848 & 0.831 & 0.897 & 0.853 & 0.866\\
\bottomrule
\end{tabular}
\end{table}

@@

\NormalTok{master\_fig }\OtherTok{\textless{}{-}} \FunctionTok{ggplot}\NormalTok{(}\AttributeTok{data =}\NormalTok{ master\_df\_long, }
                    \FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ Sample\_Size, }\AttributeTok{y =}\NormalTok{ prop, }\AttributeTok{color =}\NormalTok{ method)) }\SpecialCharTok{+} 
  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+} \FunctionTok{geom\_line}\NormalTok{() }\SpecialCharTok{+} 
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Proportion of CIs that Contain the True Parameter from Log{-}Normal Distibution"}\NormalTok{, }
       \AttributeTok{x =} \StringTok{"Sample Size"}\NormalTok{, }
       \AttributeTok{y =} \StringTok{"Proportion (of 10,000 Simulations)"}\NormalTok{, }\AttributeTok{color =} \StringTok{"Method"}\NormalTok{,}
       \AttributeTok{caption =} \StringTok{"Fig. 8: Graph showing the Accuracy of Bootstrap CI Methods for different sample sizes on Log{-}Normal Dist."}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{theme\_light}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position=}\StringTok{"bottom"}\NormalTok{)}
